\documentclass[../main/report.tex]{subfiles}
\begin{document}
\chapter{Discussion}

\subfile{../discussion/design_process.tex}

\section{PCB}
Split project schematics into several connected docs for the sake of modularity. Find some overarching hardware requirements and from there, related application notes and hardware considerations. Then some components and footprints that fit...ish with these docs.
Finally place components and route the wires between them digitally and order the PCB and components


\section{USB data input}
In the initial design, a USB connector was introduced to receive data from a host PC. 
But during the implementation phase, it was discovered to be unnecessary.
For the graphical demo, there was no need for input data from host PC. 
The MCU was able to run the program independently.

\subfile{../discussion/energy_efficiency.tex}

\section{"Jaktstart"}

Not everyone can start at the same time

\section{Redundancy design choices or w/e}
What did we end up needing?
What could we have used, that we didn't think of.

\section{Does the computer make sense?}
Is the design sane?
Were we idiots?
Have we seen the light?

\section{Can we even saturate the memory bus?}
Do you even SRAM?
Or is memory the big bottleneck

\section{SRAM bottleneck}
HDMI vs. the GPU?
Will the HDMI be fine with always being second in line?

\section{Budget}
The group was given a budget of 10000 NOK for production of the PCB and component purchases. 
We had 10 boards manufactured at the price of 10103 NOK (see appendix \cite{fig: FPGA_order} ) and ordered components at a price of 7571 NOK. (see appendix \cite{fig: component_order}) 
This made for a total cost of 17674 NOK.
The reason for going 76\% over the budget was twofold. Firstly, the PCB design took so much that time that, in addition to the size of the board, fast production and delivery from the manufacturer became necessary. 
This contributed greatly to the price.
Secondly, we bought enough components to solder at least 5 PCBs, which drove up the price a bit as well.
The total price could thus have been limited by a smaller, less backup oriented design. 
By allowing a longer production time for the manufacturer.
And finally, by not buying extra components.
\end{document}
